Assessing the impacts of time to detection distribution assumptions on detection probability estimation

### 2017-07-14 Draft.tex
Comment out unnecessary packages
Corrected 'Overbird'. Replaced .pdf's with .jpg's in Draft.tex Still need to update Supplement.tex

\documentclass[12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{parskip}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage[margin=1cm]{caption}
\usepackage{adjustbox}
\usepackage[toc,page]{appendix}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{url}
\usepackage{xr}
\usepackage{lineno}

@@

Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  
The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption can often lead to biased estimates.  
We consider a class of N-mixture models that allows for detection heterogeneity over time through a flexibly defined time-to-detection distribution (TTDD) and allows for fixed and random effects for both abundance and detection.
Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  
We specifically explore two-parameter families of TTDDs, e.g. gamma, that can additionally include a mixture component to model increased probability of detection in the initial observation period.
Based on simulation analyses, we find that modeling a TTDD by using a two-parameter family is necessary when data have a chance of arising from a distribution of this nature.
In addition, models with a mixture component can outperform non-mixture models even when the truth is non-mixture.  
Finally, we analyze an Overbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection.
We demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs but that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.

@@

\begin{figure}\centering
\includegraphics[width=\textwidth]{"Oral Prelim/AMRO_JABES"}
\caption{\label{fig:schematic} 
Illustration of fitting mixture exponential (left) and mixture gamma (right) time-to-detection distributions (TTDDs) to interval-censored removal-sampled observations (center).
The mixture TTDD consists of a continuous TTDD (thick line) plus a mixture component of first-interval detections (light gray rectangle), constituting $\gamma$ and $1-\gamma$ proportions of the population, respectively.
We estimate $\pdet$ as the proportion of the TTDD before the end of the observation period $C$, leaving an estimated proportion $1-\pdet$ undetected.
}
\end{figure}

@@

\begin{figure}\centering
\includegraphics[width=\textwidth]{"Sims/SimZero/mixture_fig"}
\caption{
Comparisons of mixture and non-mixture TTDDs.  \textit{Left}: Average posterior median detection probabilities across 100 replicate simulations.  Dashed lines show unbiased estimation.   \textit{Right}: Coverage of 90\% credible intervals across 100 replicate simulations.  Dashed lines depict a 95\% range of observed coverage that is consistent with nominal coverage.  \textit{Rows}: Data simulated from non-mixture (upper) or mixture (lower) TTDDs.  \textit{Columns}:  Data simulated from nonpeaked (left), exponential (center), or peaked (right) TTDDs.  Each inference model was fit only to datasets from the same TTDD family (e.g. lognormal to lognormal).
}
\label{fig:mixture_fig} 
\end{figure}

@@

\begin{figure}\centering
\includegraphics[width=\textwidth]{"Sims/SimZero/Posteriors_Families_draft"}
\caption{\label{fig:sim2typical} 
Representative examples of posterior distributions for $\pdet$ (left panel) and survey-level log-scale abundance (right panel).
Data are simulated from nonpeaked gamma (left column), exponential (center), and peaked gamma (right) mixture TTDDs.
True detection probabilities vary by row and are shown for comparison (dashed vertical line).
Inference models are either exponential mixture (light gray) or gamma mixture (dark gray) TTDDs.
}
\end{figure}

@@

\begin{figure}\centering
\includegraphics[width=\textwidth]{"Sims/SimZero/family_fig"}
\caption{
Comparisons of TTDDs across families.  \textit{Left}: Average posterior median detection probabilities across 100 replicate simulations.  Dashed lines show unbiased estimation.   \textit{Right}: Coverage of 90\% credible intervals across 100 replicate simulations.  Dashed lines depict a 95\% range of observed coverage that is consistent with nominal coverage.  \textit{Rows}:  Data simulated from mixture exponential, gamma, lognormal, or Weibull TTDDs.  \textit{Columns}: Data simulated from nonpeaked or exponential (left) or peaked (right) TTDDs.  All data and inference models used mixture TTDDs.
}
\label{fig:family_fig}
\end{figure}

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors}
\caption{\label{ovenposteriors} Posterior medians (black dots) with 50\% (wider line) and 95\% (narrowerline) credible intervals for the mixing parameter ($\gamma$), shape parameter ($\alpha$) as well as abundance (A) and detection (D) fixed effects and random effect standard deviations.
Posteriors are also available for the overall probability of detection and the abundance across all sites.}
\end{figure}

>>>>

\documentclass[12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
% \usepackage{amssymb}
% \usepackage{amsfonts}
% \usepackage{parskip}
\usepackage{natbib}
% \usepackage{color}
% \usepackage[dvipsnames]{xcolor}
\usepackage[margin=1cm]{caption}
\usepackage{adjustbox}
% \usepackage[toc,page]{appendix}
% \usepackage{enumitem}
% \usepackage{multirow} 
\usepackage{setspace} % for \doublespacing
% \usepackage{url}
\usepackage{xr} % for cross referencing to supplement
% \usepackage{lineno}

@@

Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  
The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption can often lead to biased estimates.  
We consider a class of N-mixture models that allows for detection heterogeneity over time through a flexibly defined time-to-detection distribution (TTDD) and allows for fixed and random effects for both abundance and detection.
Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  
We specifically explore two-parameter families of TTDDs, e.g. gamma, that can additionally include a mixture component to model increased probability of detection in the initial observation period.
Based on simulation analyses, we find that modeling a TTDD by using a two-parameter family is necessary when data have a chance of arising from a distribution of this nature.
In addition, models with a mixture component can outperform non-mixture models even when the truth is non-mixture.  
Finally, we analyze an Ovenbird data set from the Chippewa National Forest using mixed effect models for both abundance and detection.
We demonstrate that the effects of explanatory variables on abundance and detection are consistent across mixture TTDDs but that flexible TTDDs result in lower estimated probabilities of detection and therefore higher estimates of abundance.

@@

\begin{figure}\centering
\includegraphics[width=\textwidth]{"Oral Prelim/AMRO_JABES"} % Source file: Oral Prelim/TTDD_Demo.R
\caption{\label{fig:schematic} 
Illustration of fitting mixture exponential (left) and mixture gamma (right) time-to-detection distributions (TTDDs) to interval-censored removal-sampled observations (center).
The mixture TTDD consists of a continuous TTDD (thick line) plus a mixture component of first-interval detections (light gray rectangle), constituting $\gamma$ and $1-\gamma$ proportions of the population, respectively.
We estimate $\pdet$ as the proportion of the TTDD before the end of the observation period $C$, leaving an estimated proportion $1-\pdet$ undetected.
}
\end{figure}

@@

\begin{figure}\centering
\includegraphics[width=\textwidth]{"Sims/SimZero/mixture_fig"} % Source file: Sims/SimZero/Outputs_Code_zero.R
\caption{
Comparisons of mixture and non-mixture TTDDs.  \textit{Left}: Average posterior median detection probabilities across 100 replicate simulations.  Dashed lines show unbiased estimation.   \textit{Right}: Coverage of 90\% credible intervals across 100 replicate simulations.  Dashed lines depict a 95\% range of observed coverage that is consistent with nominal coverage.  \textit{Rows}: Data simulated from non-mixture (upper) or mixture (lower) TTDDs.  \textit{Columns}:  Data simulated from nonpeaked (left), exponential (center), or peaked (right) TTDDs.  Each inference model was fit only to datasets from the same TTDD family (e.g. lognormal to lognormal).
}
\label{fig:mixture_fig} 
\end{figure}

@@

\begin{figure}\centering
\includegraphics[width=\textwidth]{"Sims/SimZero/Posteriors_Families_draft"} % Source file: Sims/SimZero/RepresentativePosteriors.R
\caption{\label{fig:sim2typical} 
Representative examples of posterior distributions for $\pdet$ (left panel) and survey-level log-scale abundance (right panel).
Data are simulated from nonpeaked gamma (left column), exponential (center), and peaked gamma (right) mixture TTDDs.
True detection probabilities vary by row and are shown for comparison (dashed vertical line).
Inference models are either exponential mixture (light gray) or gamma mixture (dark gray) TTDDs.
}
\end{figure}

@@

\begin{figure}\centering
\includegraphics[width=\textwidth]{"Sims/SimZero/family_fig"} % Source file: Sims/SimZero/Outputs_Code_zero.R
\caption{
Comparisons of TTDDs across families.  \textit{Left}: Average posterior median detection probabilities across 100 replicate simulations.  Dashed lines show unbiased estimation.   \textit{Right}: Coverage of 90\% credible intervals across 100 replicate simulations.  Dashed lines depict a 95\% range of observed coverage that is consistent with nominal coverage.  \textit{Rows}:  Data simulated from mixture exponential, gamma, lognormal, or Weibull TTDDs.  \textit{Columns}: Data simulated from nonpeaked or exponential (left) or peaked (right) TTDDs.  All data and inference models used mixture TTDDs.
}
\label{fig:family_fig}
\end{figure}

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.95\textwidth]{OVEN/oven_sum/OVEN_Posteriors} % Source file: OVEN/Compile.R
\caption{\label{ovenposteriors} Posterior medians (black dots) with 50\% (wider line) and 95\% (narrowerline) credible intervals for the mixing parameter ($\gamma$), shape parameter ($\alpha$) as well as abundance (A) and detection (D) fixed effects and random effect standard deviations.
Posteriors are also available for the overall probability of detection and the abundance across all sites.}
\end{figure}
